Tight Bounds for Dynamic Balanced Graph Partitioning

### 2020-05-11 paper/disc20clustering.tex
comments

A~\emph{ground set} is a group of nodes that repeatedly communicate when an algorithm splits them (places at different clusters).
Their role is to ensure that the algorithm maintains a~partition of ground sets into clusters.
We emphasize that ground sets are revealed only upon their nodes are split.
The adversary constructs ground sets depending on the choices of a deterministic online algorithm.
Once we construct a ground set, it lasts until the end of the input sequence.
We say that a ground set is a \emph{singleton} if it contains exactly one node.
We say that the node that belongs to a singleton ground set is \emph{isolated}.

@@

\noindent
\textbf{Resource augmentation.}
The majority of work on Online Balanced Partitioning so far \cite{repartition-disc,sigmetrics19_partitioning} focuses on the scenario with resource augmentation, where the clusters of an online algorithm are larger than the clusters of the offline optimal algorithm that we compare the performance to.
We can adjust our construction to show a lower bound of $\Omega(\ell)$ for resource augmentation.

@@

We present an asymptotically optimal algorithm for the learning problem.
The algorithm  collocates  a pair as soon as they communicate and it never separates them.
In order to preserve collocated pairs,
we employ the concept of components,
introduced by Avin et al. \cite{repartition-disc}.

%\medskip
%In our algorithms, 
A~\emph{component} is a subset of frequently communicating nodes.
We keep all nodes of a component always collocated in the same cluster,
i.e., when we  move a~node,
we move the whole component that contains it.
A~partition that maintains this invariant is a \emph{component respecting} partition.
%We say that a~partition is \emph{component respecting}
%if the nodes belonging to the same component are always collocated.
We say that a component is \emph{singleton} if it contains exactly one node,
and the node in this component is \emph{isolated}.

@@

\noindent
\textbf{Perfect Partition Learner algorithm.}
Fix the initial partition
$P_I :=\{ I_1, \dots, I_{\ell}\}$ and \OPT's final partition
$P_F := \{F_1, \dots, F_{\ell}\}$.
The \emph{distance} of a partition
 $P = \{C_1, \dots, C_{\ell}\}$ from the initial partition,
defined as 
$\Delta(P, P_I) := \sum_{j=1}^{\ell} | C_j \setminus I_j |$,
 is the number of nodes in $P$ that do not reside in their initial cluster.
In other words,
at least $\Delta(P, P_I)$ node migrations are required in order to reach the partition $P$ from $P_I$, and thus
$\OPT \geq \Delta^*:= \Delta(P_F, P_I) $.
We  use $ \Delta(P_F) $ wherever the initial partition is clear from the context.
The scheme of the algorithm can be found in the  appendix (Algorithm \ref{alg:ppl}).

>>>>

A~\emph{ground set} is a group of nodes that repeatedly communicate when an algorithm splits them (places at different clusters).
Their role is to ensure that the algorithm maintains a~partition of ground sets into clusters.
We emphasize that ground sets are revealed only upon their nodes are split.
The adversary constructs ground sets depending on the choices of a deterministic online algorithm.
Once we construct a ground set, it lasts until the end of the algorithm's execution.
We say that a ground set is a \emph{singleton} if it contains exactly one node.
We say that the node that belongs to a singleton ground set is \emph{isolated}.

@@

\noindent
\textbf{Resource augmentation.}
The majority of work on Online Balanced Partitioning so far \cite{repartition-disc,sigmetrics19_partitioning} focuses on the scenario with resource augmentation, where the clusters of an online algorithm are larger than the clusters of the offline optimal algorithm that we compare the performance to.
We can adjust our construction to show a lower bound of $\Omega(\ell)$ for a setting with resource augmentation.

@@

We present an asymptotically optimal algorithm for the learning problem.
The algorithm  collocates  a pair as soon as they communicate and it never separates them.
In order to preserve collocated pairs,
we employ the concept of components,
introduced by Avin~et~al.~\cite{repartition-disc}.

%\medskip
%In our algorithms, 
A~\emph{component} is a subset of frequently communicating nodes \maciek{this is not a good definition. intuitions, maybe}.
We keep all nodes of a component always collocated in the same cluster,
i.e., when we  move a~node,
we move the whole component that contains it.
A~partition that maintains this invariant is a \emph{component respecting} partition.
%We say that a~partition is \emph{component respecting}
%if the nodes belonging to the same component are always collocated.
We say that a component is \emph{singleton} if it contains exactly one node,
and the node in this component is \emph{isolated}.

@@

\noindent
\textbf{Perfect Partition Learner algorithm.}
Fix the initial partition
$P_I :=\{ I_1, \dots, I_{\ell}\}$ and \OPT's final partition
$P_F := \{F_1, \dots, F_{\ell}\}$.
The \emph{distance} of a partition
 $P = \{C_1, \dots, C_{\ell}\}$ from the initial partition,
defined as 
$\Delta(P, P_I) := \sum_{j=1}^{\ell} | C_j \setminus I_j |$,
 is the number of nodes in $P$ that do not reside in their initial cluster.
In other words,
at least $\Delta(P, P_I)$ node migrations are required in order to reach the partition $P$ from $P_I$, and thus
$\OPT \geq \Delta^*:= \Delta(P_F, P_I) $.
We  use $ \Delta(P_F) $ wherever the initial partition is clear from the context \maciek{It always is, we fixed it in the beginning. Maybe just define it as the distance from initial partition, but don't pass it as an argument}.
The scheme of the algorithm can be found in the  appendix (Algorithm \ref{alg:ppl}).
